Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2012-10-16 paper.tex
minor

\begin{lemma}\label{lem:supDLlambdan}
With $N$ a finite constant $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
\end{lemma}
\begin{IEEEproof}
Put
\[
Z_i(\lambda) = \begin{cases}
R_i\cos(\lambda + \Phi_i) , & i \in P \\
R_i\cos\sfracpart{\lambda + \Phi_i} , & i \in D
\end{cases}
\]
so that
\[
D_L(\lambda) = G_L(\lambda) - T_L(\lambda) = \frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
\]
Now $Z_1(\lambda_n), \dots,Z_L(\lambda_n)$ are independent with finite variance (because $\expect R_i^2$ is finite), so for each $n =1, \dots, N$,
\[
\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda_n) - \expect Z_i(\lambda)\big)} \rightarrow 0
\]
almost surely as $L\rightarrow\infty$ by the Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Since $N$ is a constant,
\[
U_L = \sup_{n=1,\dots,N}\sabs{D_L(\lambda_n)} \leq \sum_{n=1}^N \sabs{D_L(\lambda_n)} \rightarrow 0
\]
almost surely to zero as $L \rightarrow \infty$.
\end{IEEEproof}

@@

\begin{lemma}~\label{lem:diffatlambdaL}
The derivative of $G_L$ exists, and is equal to zero, at $\hat{\lambda}_L$.  That is,
\[
G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0.
\]
\end{lemma}
\begin{IEEEproof}
Observe that 
\[
G_L(\lambda) = \frac{1}{L}\sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}
\] 
is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ with $R_i > 0$.  Let $q_i$ be the smallest number from the interval $[-\tfrac{\pi}{M}, 0]$ such that
\[
 \sin(q_i) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}
\]
where $\eta = e^{-j2\pi/M} - 1$.  Observe that if $R_i > 0$ then $q_i < 0$.  Lemma~\ref{lem:fracpartlambdahatnotpi} shows that,
\[
\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i
\]
for all $i \in D$.  Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ for $i \in D$ when $R_i > 0$ and therefore $G_L$ is differentiable at $\hat{\lambda}_L$.

It remains to show that $G_L^\prime(\hat{\lambda}_L) = 0$.  If $\sabs{G_L(\hat{\lambda}_L)} > 0$ then $\sabs{G_L(\lambda)}$ is differentiable at $\hat{\lambda}_L$ and, since $\hat{\lambda}_L$ is a maximiser of $\sabs{G_L(\lambda)}$,
\[
0 = \frac{d \sabs{G_L(\lambda)}}{d \lambda}(\hat{\lambda}_L) = \begin{cases} 
G_L^\prime(\hat{\lambda}_L), & G_L(\hat{\lambda}_L) > 0\\
-G_L^\prime(\hat{\lambda}_L), & G_L(\hat{\lambda}_L) < 0.
\end{cases}
\]
Thus, $G_L^\prime(\hat{\lambda}_L) = 0$ if $\sabs{G_L(\hat{\lambda}_L)} > 0$.  On the other hand $\sabs{G_L(\hat{\lambda}_L)} = 0$ only if $G_L(\lambda) = 0$ for all $\lambda \in \reals$, and in this case $G_L^\prime(\hat{\lambda}_L) = 0$ also. 
\end{IEEEproof}

\begin{lemma}\label{lem:fracpartlambdahatnotpi} Let $q_i, i \in D$ be defined as in Lemma~\ref{lem:diffatlambdaL}.  Then $\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i$ for all $i \in D$.
\end{lemma}
\begin{IEEEproof}
Recall that $\{\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D\}$ defined in \eqref{eq:hatdfinxtheta} are the minimisers of the function 
\[
SS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2,
\]
defined in~\eqref{eq:SSdatasymbols}. The proof now proceeds by contradiction.  Assume that 
\begin{equation}\label{eq:fraclambassumption}
\sfracpart{\hat{\lambda}_L + \Phi_k} > \frac{\pi}{M} + q_k
\end{equation}
for some $k \in D$.  Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$.  We will show that 
\[
SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\}),
\]
violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  First observe that,
\begin{align*}
Y(\{ r_i, i \in D\}) = \sum_{i \in P} y_ip_i^* + \sum_{i \in D} y_i\hat{r}_i^* = \hat{Y} + \eta y_k\hat{d}_k^*,
\end{align*}
where $\eta = e^{-j2\pi/M} - 1$ and $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
\begin{align*}
SS&(\{r_i, i \in D\}) \\
&= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
&= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
&= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
&= SS(\{\hat{d}_i, i \in D\}) - D,
\end{align*}
where 
\[
D = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\sabs{ \eta y_k }^2.
\]
Now $\frac{1}{L}\hat{Y} = \hat{a} = \hat{\rho} e^{j\hat{\theta}}$ from~\eqref{eq:hata} and using~\eqref{eq:yethetadhat},
\[
\frac{1}{L} \hat{Y}^* y_k \hat{d}_k^* = \hat{\rho} \rho_0 R_k e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}},
\]
so that
\begin{equation}\label{eq:Ddefn}
D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\sabs{\eta}^2\rho_0^2 R_k^2.
\end{equation}
Let $v = \sfracpart{\hat{\lambda}_L + \Phi_k} - \tfrac{\pi}{M}$ so that
\begin{align*}
\eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}} &= (e^{-j2\pi/M} - 1) e^{j \pi/M} e^{jv} \\
&= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
&= -2 \sin(\tfrac{\pi}{M}) e^{j ( v + \pi/2) },
\end{align*}
and
\begin{align*}
\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) &= -2 \sin(\tfrac{\pi}{M}) \cos(v + \tfrac{\pi}{2}) \\
 &= 2 \sin(\tfrac{\pi}{M}) \sin(v),
\end{align*}
Because we assumed~\eqref{eq:fraclambassumption}, it follows that $0 > v > q_k$ and, from the definition of $q_k$,
\[
0 > \sin(v) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}.
\]
Substituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  Thus,~\eqref{eq:fraclambassumption} is false.

>>>>

\begin{lemma}\label{lem:supDLlambdan}
With $N$ a finite constant $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
\end{lemma}
\begin{IEEEproof}
Put
\[
Z_i(\lambda) = \begin{cases}
R_i\cos(\lambda + \Phi_i) , & i \in P \\
R_i\cos\sfracpart{\lambda + \Phi_i} , & i \in D
\end{cases}
\]
so that
\[
D_L(\lambda) = G_L(\lambda) - T_L(\lambda) = \frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
\]
Now $Z_1(\lambda_n), \dots,Z_L(\lambda_n)$ are independent with finite variance (because $\expect R_i^2$ is finite), so for each $n =1, \dots, N$,
\[
\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda_n) - \expect Z_i(\lambda)\big)} \rightarrow 0
\]
almost surely as $L\rightarrow\infty$ by Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Since $N$ is a constant,
\[
U_L = \sup_{n=1,\dots,N}\sabs{D_L(\lambda_n)} \leq \sum_{n=1}^N \sabs{D_L(\lambda_n)} \rightarrow 0
\]
almost surely to zero as $L \rightarrow \infty$.
\end{IEEEproof}

@@

\begin{lemma}~\label{lem:diffatlambdaL}
The derivative of $G_L$ exists, and is equal to zero, at $\hat{\lambda}_L$.  That is,
\[
G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0.
\]
\end{lemma}
\begin{IEEEproof}
Observe that 
\[
G_L(\lambda) = \frac{1}{L}\sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}
\] 
is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ with $R_i > 0$.  Let $q_i$ be the smallest number from the interval $[-\tfrac{\pi}{M}, 0]$ such that
\[
 \sin(q_i) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}
\]
where $\eta = e^{-j2\pi/M} - 1$.  Observe that if $R_i > 0$ then $q_i < 0$.  Lemma~\ref{lem:fracpartlambdahatnotpi} shows that,
\[
\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i
\]
for all $i \in D$.  Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ for $i \in D$ when $R_i > 0$ and therefore $G_L$ is differentiable at $\hat{\lambda}_L$.  That $G_L^\prime(\hat{\lambda}_L) = 0$ follows since $\hat{\lambda}_L$ is a maximiser of $G_L$.
\end{IEEEproof}

\begin{lemma}\label{lem:fracpartlambdahatnotpi} Let $q_i$ be defined as in Lemma~\ref{lem:diffatlambdaL}.  Then $\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i$ for all $i \in D$.
\end{lemma}
\begin{IEEEproof}
Recall that $\{\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D\}$ defined in \eqref{eq:hatdfinxtheta} are the minimisers of the function 
\[
SS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2,
\]
defined in~\eqref{eq:SSdatasymbols}. The proof now proceeds by contradiction.  Assume that 
\begin{equation}\label{eq:fraclambassumption}
\sfracpart{\hat{\lambda}_L + \Phi_k} > \frac{\pi}{M} + q_k
\end{equation}
for some $k \in D$.  Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$.  We will show that 
\[
SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\}),
\]
violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  First observe that,
\begin{align*}
Y(\{ r_i, i \in D\}) = \sum_{i \in P} y_ip_i^* + \sum_{i \in D} y_i\hat{r}_i^* = \hat{Y} + \eta y_k\hat{d}_k^*,
\end{align*}
where $\eta = e^{-j2\pi/M} - 1$ and $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
\begin{align*}
SS&(\{r_i, i \in D\}) \\
&= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
&= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
&= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
&= SS(\{\hat{d}_i, i \in D\}) - D,
\end{align*}
where 
\[
D = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\sabs{ \eta y_k }^2.
\]
Now $\frac{1}{L}\hat{Y} = \hat{a} = \hat{\rho} e^{j\hat{\theta}}$ from~\eqref{eq:hata} and using~\eqref{eq:yethetadhat},
\[
\frac{1}{L} \hat{Y}^* y_k \hat{d}_k^* = \hat{\rho} \rho_0 R_k e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}},
\]
so that
\begin{equation}\label{eq:Ddefn}
D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\sabs{\eta}^2\rho_0^2 R_k^2.
\end{equation}
Let $v = \sfracpart{\hat{\lambda}_L + \Phi_k} - \tfrac{\pi}{M}$ so that
\begin{align*}
\eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}} &= (e^{-j2\pi/M} - 1) e^{j \pi/M} e^{jv} \\
&= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
&= -2 \sin(\tfrac{\pi}{M}) e^{j ( v + \pi/2) },
\end{align*}
and
\begin{align*}
\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) &= -2 \sin(\tfrac{\pi}{M}) \cos(v + \tfrac{\pi}{2}) \\
 &= 2 \sin(\tfrac{\pi}{M}) \sin(v),
\end{align*}
Because we assumed~\eqref{eq:fraclambassumption}, it follows that $0 > v > q_k$ and, from the definition of $q_k$,
\[
0 > \sin(v) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}.
\]
Substituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  Thus,~\eqref{eq:fraclambassumption} is false.
